# 2024-10-03 - Analyze Linear Chain Properties

```{r}
library(tidyverse)
library(cowplot)
library(RColorBrewer)
library(khroma)
library(rstatix)
library(knitr)
library(kableExtra)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

```{r}
experiment_slug <- "2024-10-03-linear-chain"
working_directory <- paste(
  "experiments",
  experiment_slug,
  "analysis",
  sep = "/"
)
# Adjust working directory if being knitted for bookdown build.
if (exists("bookdown")) {
  if (bookdown) {
    working_directory <- paste0(
      bookdown_wd_prefix,
      working_directory
    )
  }
}
```

```{r}
# Configure our default graphing theme
theme_set(theme_cowplot())
# Create a directory to store plots
plot_dir <- paste(
  working_directory,
  "plots",
  sep = "/"
)

dir.create(
  plot_dir,
  showWarnings = FALSE
)
```

Load graph birth data

```{r}
graph_birth_data_path <- paste(
  working_directory,
  "data",
  "graph_birth_info.csv",
  sep = "/"
)

graph_birth_data <- read_csv(graph_birth_data_path)
```

Load node info data

```{r}
node_info_data_path <- paste(
  working_directory,
  "data",
  "node_info_linear-chain.csv",
  sep = "/"
)

node_info_data <- read_csv(node_info_data_path)
```

## Frequency of birth counts

```{r}
graph_birth_plt <- graph_birth_data %>%
  ggplot(
    aes(x = births)
  ) +
  scale_x_continuous(
    name = "Location birth count"
  ) +
  geom_density(
    fill="#69b3a2", color="#e9ecef", alpha=0.8
  )

ggsave(
  filename = paste0(plot_dir, "/graph_births.pdf"),
  plot = graph_birth_plt,
  width = 15,
  height = 10
)
```

```{r}
graph_birth_log_plt <- graph_birth_data %>%
  ggplot(
    aes(x = births)
  ) +
  scale_x_log10(
    name = "Location birth count (log scale)"
  ) +
  geom_density(
    fill="#69b3a2", color="#e9ecef", alpha=0.8
  )

ggsave(
  filename = paste0(plot_dir, "/graph_births_log.pdf"),
  plot = graph_birth_log_plt,
  width = 15,
  height = 10
)
```

## Proportion of task appearances by id

```{r}
task_appearances_by_id_plt <- node_info_data %>%
  ggplot(
    aes(
      x = loc_id,
      y = all_task_prop
    )
  ) +
  geom_bar(stat = "identity")

ggsave(
  filename = paste0(plot_dir, "/task_app_by_id.pdf"),
  plot = task_appearances_by_id_plt,
  width = 20,
  height = 10
)
```

## Proportion of actual births by id

```{r}
act_births_by_id_plt <- node_info_data %>%
  ggplot(
    aes(
      x = loc_id,
      y = actual_births_prop
    )
  ) +
  geom_bar(stat = "identity")

ggsave(
  filename = paste0(plot_dir, "/act_births_by_id.pdf"),
  plot = act_births_by_id_plt,
  width = 20,
  height = 10
)
```

## Proportion of expected births by id

```{r}
expected_births_by_id_plt <- node_info_data %>%
  ggplot(
    aes(
      x = loc_id,
      y = expected_births_prop
    )
  ) +
  geom_bar(stat = "identity")

ggsave(
  filename = paste0(plot_dir, "/expected_births_by_id.pdf"),
  plot = expected_births_by_id_plt,
  width = 20,
  height = 10
)
```

## Task appearances frequency distribution

```{r}
longer_props <- node_info_data %>% pivot_longer(
  cols = c(all_task_prop, expected_births_prop, actual_births_prop),
  names_to = "proportion_type",
  values_to = "proportion"
)

task_appearances_freq_dist_plt <- longer_props %>%
  ggplot(
    aes(
      x = proportion,
      group = proportion_type,
      fill = proportion_type
    )
  ) +
  geom_density(alpha=0.4)

ggsave(
  filename = paste0(plot_dir, "/all_task_prop_freq_dist_plt.pdf"),
  plot = task_appearances_freq_dist_plt,
  width = 20,
  height = 10
)
```

```{r}
longer_props <- node_info_data %>% pivot_longer(
  cols = c(all_task_prop, expected_births_prop, actual_births_prop),
  names_to = "proportion_type",
  values_to = "proportion"
)

task_appearances_freq_dist_plt <- longer_props %>%
  filter(proportion_type != "expected_births_prop") %>%
  ggplot(
    aes(
      x = proportion,
      group = proportion_type,
      fill = proportion_type
    )
  ) +
  geom_density(alpha=0.4)

ggsave(
  filename = paste0(plot_dir, "/all_task_prop_freq_dist_plt2.pdf"),
  plot = task_appearances_freq_dist_plt,
  width = 20,
  height = 10
)
```

## All task prop vs expected

```{r}
all_task_prop_vs_expected_births_plt <- node_info_data %>%
  ggplot(
    aes(
      x = all_task_prop_vs_expected_births
    )
  ) +
  geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8) +
  geom_vline(xintercept = 0, color = "black")

ggsave(
  filename = paste0(plot_dir, "/all_task_prop_vs_expected_births_plt.pdf"),
  plot = all_task_prop_vs_expected_births_plt,
  width = 20,
  height = 10
)
```

## All task prop vs actual

```{r}
all_task_prop_vs_actual_births_plt <- node_info_data %>%
  ggplot(
    aes(
      x = all_task_prop_vs_actual_births
    )
  ) +
  geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8) +
  geom_vline(xintercept = 0, color = "black")

ggsave(
  filename = paste0(plot_dir, "/all_task_prop_vs_actual_births_plt.pdf"),
  plot = all_task_prop_vs_actual_births_plt,
  width = 20,
  height = 10
)
```

## All task prop vs ?? (correlation)

```{r}
all_task_prop_vs_actual_births_cor_plt <- node_info_data %>%
  ggplot(
    aes(
      x = actual_births_prop,
      y = all_task_prop
    )
  ) +
  geom_point() +
  geom_smooth(method = "lm", formula = y~x)


ggsave(
  filename = paste0(plot_dir, "/all_task_prop_vs_actual_births_cor.pdf"),
  plot = all_task_prop_vs_actual_births_cor_plt,
  width = 20,
  height = 10
)
```

```{r}
all_task_prop_vs_expected_births_cor_plt <- node_info_data %>%
  ggplot(
    aes(
      x = expected_births_prop,
      y = all_task_prop
    )
  ) +
  geom_point() +
  geom_smooth(method = "lm", formula = y~x)


ggsave(
  filename = paste0(plot_dir, "/all_task_prop_vs_expected_births_cor_plt.pdf"),
  plot = all_task_prop_vs_expected_births_cor_plt,
  width = 20,
  height = 10
)
```

```{r}
property_data_long <- node_info_data %>%
  pivot_longer(
    cols = c(
      percolation_centrality,
      harmonic_centrality,
      subgraph_centrality,
      eigenvector_centrality,
      load_centrality,
      second_order_centrality,
      triangles,
      closeness_centrality,
      information_centrality,
      clustering,
      degree,
      actual_births_prop,
      expected_births_prop
    ),
    names_to = "node_property",
    values_to = "node_property_value"
  )

plt <- property_data_long %>%
  ggplot(
    aes(
      x = node_property_value,
      y = all_task_prop
    )
  ) +
  geom_point() +
  geom_smooth(method = "lm", formula = y~x) +
  facet_wrap(~node_property, scales = "free")

ggsave(
  filename = paste0(plot_dir, "/node_properties_cor.pdf"),
  plot = plt,
  width = 20,
  height = 20
)
```

## Outliers in task appearance proportion

```{r}
iqr <- IQR(node_info_data$all_task_prop)
upper_bound <- quantile(node_info_data$all_task_prop, probs=c( .75), na.rm = FALSE)+1.5*iqr
lower_bound <- quantile(node_info_data$all_task_prop, probs=c( .75), na.rm = FALSE)-1.5*iqr

greater_outliers <- node_info_data %>%
  filter(all_task_prop >= upper_bound) %>%
  arrange(
    # desc(abs(all_task_prop)),
    loc_id
  ) %>%
  select(
    all_task_prop, loc_id
  )
print(greater_outliers, n = 100)

lower_outliers <- node_info_data %>%
  filter(all_task_prop <= lower_bound) %>%
  arrange(
    # desc(abs(all_task_prop)),
    loc_id
  ) %>%
  select(
    all_task_prop, loc_id
  )
print(lower_outliers, n = 350)

mean(node_info_data$all_task_prop)
```


