# Manuscript figures

## Analysis setup

```{r}
library(tidyverse)
library(cowplot)
library(RColorBrewer)
library(khroma)
library(rstatix)
library(knitr)
library(kableExtra)
library(boot)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

```{r}
varied_graph_experiment_slug <- "analysis_of_exploratory_data"
varied_graph_working_directory <- paste(
  "third-party",
  "evo_spatial_discoveries",
  "experiments",
  varied_graph_experiment_slug,
  sep = "/"
)

toroid_graph_experiment_slug <- "analysis_of_exploratory_toroid_data"
toroid_graph_working_directory <- paste(
  "third-party",
  "evo_spatial_discoveries",
  "experiments",
  toroid_graph_experiment_slug,
  sep = "/"
)

plt_dump_dir <- paste(
  "experiments",
  "mabe2-exps",
  "analyses",
  "plots",
  sep = "/"
)
```

```{r}
# Configure our default graphing theme
theme_set(theme_cowplot())
# Create a directory to store plots
plot_dir <- plt_dump_dir

dir.create(
  plot_dir,
  showWarnings = FALSE
)
```

## Load data

General parameters

```{r}
max_generation <- 100000
```

Load varied graph max organism data

```{r}
varied_graph_max_org_data_path <- paste(
  varied_graph_working_directory,
  "data",
  "combined_max_org_data.csv",
  sep = "/"
)
# Data file has time series
varied_graph_max_org_data_ts <- read_csv(varied_graph_max_org_data_path)
varied_graph_max_org_data_ts <- varied_graph_max_org_data_ts %>%
  mutate(
    landscape = as.factor(landscape),
    structure = as.factor(structure),
    experiment_set = "varied-graph"
  ) %>%
  mutate(
    valleys_crossed = case_when(
      landscape == "Valley crossing" ~ round(log(fitness, base = 1.5)),
      .default = 0
    )
  )
# Get tibble with just final generation
varied_graph_max_org_data <- varied_graph_max_org_data_ts %>%
  filter(generation == max_generation)
```

Load toroid max organism data

```{r}
toroid_graph_max_org_data_path <- paste(
  toroid_graph_working_directory,
  "data",
  "combined_max_org_data.csv",
  sep = "/"
)

# Data file has time series
toroid_graph_max_org_data_ts <- read_csv(toroid_graph_max_org_data_path)
toroid_graph_max_org_data_ts <- toroid_graph_max_org_data_ts %>%
  mutate(
    landscape = as.factor(landscape),
    structure = as.factor(structure),
    experiment_set = "squished-toroid"
  ) %>%
  mutate(
    valleys_crossed = case_when(
      landscape == "Valley crossing" ~ round(log(fitness, base = 1.5)),
      .default = 0
    )
  )
# Get tibble with just final generation
toroid_graph_max_org_data <- toroid_graph_max_org_data_ts %>%
  filter(generation == max_generation)
```

Combine toroid and varied graph data into single data frame.

```{r}
max_org_data_ts <- full_join(
  varied_graph_max_org_data_ts,
  toroid_graph_max_org_data_ts
)
max_org_data_ts <- max_org_data_ts %>%
  mutate(
    experiment_set = as.factor(experiment_set)
  )

structure_label_map <- list(
  "1_3600" = "1x3600",
  "15_240" = "15x240",
  "2_1800" = "2x1800",
  "3_1200" = "3x1200",
  "30_120" = "30x120",
  "4_900" = "4x900",
  "60_60" = "60x60",
  "well_mixed" = "Well-mixed",
  "comet_kite" = "Comet-kite",
  "windmill" = "Windmill",
  "clique_ring" = "Clique ring",
  "cycle" = "Cycle",
  "linear_chain" = "Linear chain",
  "lattice" = "Lattice",
  "random_waxman" = "Random Waxman",
  "star" = "Star",
  "wheel" = "Wheel"
)
toroid_exp_structs <- c(
  "1x3600",
  "2x1800",
  "3x1200",
  "4x900",
  "15x240",
  "30x120",
  "60x60"
)
varied_graph_exp_structs <- c(
  "Well-mixed",
  "Comet-kite",
  "Windmill",
  "Clique ring",
  "Cycle",
  "Linear chain",
  "Lattice",
  "Random Waxman",
  "Star",
  "Wheel"
)

structure_levels <- c(
  "1_3600",
  "2_1800",
  "3_1200",
  "4_900",
  "15_240",
  "30_120",
  "60_60",
  "well_mixed",
  "comet_kite",
  "windmill",
  "clique_ring",
  "cycle",
  "linear_chain",
  "lattice",
  "random_waxman",
  "star",
  "wheel"
)

structure_labels <- c()
for (struct_id in structure_levels) {
  structure_labels <- append(structure_labels, structure_label_map[[struct_id]])
}

max_org_data_ts <- max_org_data_ts %>%
  mutate(
    structure = factor(
      structure,
      levels = structure_levels,
      labels = structure_labels
    )
  )
# Get tibble with just final generation
max_org_data <- max_org_data_ts %>%
  filter(generation == max_generation)
```

Check that replicate count for each condition matches expectations.

```{r}
run_summary <- max_org_data %>%
  group_by(experiment_set, landscape, structure) %>%
  summarize(
    n = n()
  )
print(run_summary, n = 60)
```

Calculate time to max fitness for single-gradient landscape.

```{r}
# Find all rows with maximum fitness value, then get row with minimum generation,
#  then project out expected generation to max (for runs that didn't finish)
max_possible_fit = 50
time_to_max_single_gradient <- max_org_data_ts %>%
  filter(landscape == "Single gradient") %>%
  group_by(rep, structure, experiment_set) %>%
  slice_max(
    fitness,
    n = 1
  ) %>%
  slice_min(
    generation,
    n = 1
  ) %>%
  mutate(
    proj_gen_max = (max_possible_fit / fitness) * generation
  )
```

Calculate bootstrapped 99% confidence intervals for projected generation to maximum
fitness for single-gradient runs.

```{r}
# Define sample mean function
samplemean <- function(x, d) {
  return(mean(x[d]))
}

summary_gen_to_max <- tibble(
  experiment_set = character(),
  structure = character(),
  proj_gen_max_mean = double(),
  proj_gen_max_mean_ci_low = double(),
  proj_gen_max_mean_ci_high = double()
)
structures <- levels(time_to_max_single_gradient$structure)
experiment_sets <- levels(time_to_max_single_gradient$experiment_set)
for (exp_set in experiment_sets) {
  # cat("--------", exp_set, "--------")
  for (struct in structures) {
    # print(struct)
    s_data <- filter(
      time_to_max_single_gradient,
      (structure == struct) & (experiment_set == exp_set)
    )$proj_gen_max
    if (length(s_data) == 0) {
      next
    }
    boot_result <- boot(
      s_data,
      statistic = samplemean,
      R = 10000
    )
    result_ci <- boot.ci(boot_result, conf = 0.99, type = "perc")
    m <- result_ci$t0
    low <- result_ci$percent[4]
    high <- result_ci$percent[5]

    summary_gen_to_max <- summary_gen_to_max %>%
      add_row(
        experiment_set = exp_set,
        structure = struct,
        proj_gen_max_mean = m,
        proj_gen_max_mean_ci_low = low,
        proj_gen_max_mean_ci_high = high
      )
  }
}
summary_gen_to_max <- summary_gen_to_max %>%
  mutate(
    experiment_set = as.factor(experiment_set),
    structure = factor(
      structure,
      levels = structure_labels
    )
  )
```

## Varied graph type experiment set

### Smooth gradient - Time to maximum fitness

```{r}

# Identify median for well-mixed
wm_median <- median(
  filter(
    time_to_max_single_gradient,
    (structure == "Well-mixed") & (experiment_set == "varied-graph")
  )$proj_gen_max
)

vg_clr_pal <- scale_color_discreterainbow()$palette(
  length(varied_graph_exp_structs)
)

# color("discreterainbow")(length(varied_graph_exp_structs))
wm_color <- vg_clr_pal[1]
hline_width <- 0.66
vg_smooth_grad_time_to_max_plt <- summary_gen_to_max %>%
  filter(experiment_set == "varied-graph") %>%
  ggplot(
    mapping = aes(
      x = structure,
      y = proj_gen_max_mean,
      fill = structure,
      color = structure
    )
  ) +
  # geom_point() +
  geom_col() +
  geom_linerange(
    aes(
      ymin = proj_gen_max_mean_ci_low,
      ymax = proj_gen_max_mean_ci_high
    ),
    color = "black",
    linewidth = 0.75,
    lineend = "round"
  ) +
  geom_hline(
    yintercept = max_generation,
    linetype = "solid",
    linewidth = 1,
    color = "darkgrey",
    alpha = hline_width
  ) +
  geom_hline(
    yintercept = wm_median,
    color = wm_color,
    linetype = "dashed",
    linewidth = hline_width
  ) +
  scale_x_discrete(
    breaks = rev(varied_graph_exp_structs),
    limits = rev(varied_graph_exp_structs)
  ) +
  scale_color_manual(
    values = vg_clr_pal
  ) +
  scale_fill_manual(
    values = vg_clr_pal
  ) +
  coord_flip() +
  theme(
    legend.position = "none"
  )

vg_smooth_grad_time_to_max_plt
```

### Multi-path exploration - Maximum fitness

```{r}

mp_wm_median <- median(
  filter(
    max_org_data,
    (landscape == "Multipath") & (structure == "Well-mixed") & (experiment_set == "varied-graph")
  )$fitness
)

vg_multipath_fitness_plt <- max_org_data %>%
  filter(experiment_set == "varied-graph" & landscape == "Multipath") %>%
  mutate(
    structure = factor(
      structure,
      levels = varied_graph_exp_structs,
      labels = varied_graph_exp_structs
    )
  ) %>%
  ggplot(
    mapping = aes(
      x = structure,
      y = fitness,
      fill = structure
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping = aes(color = structure),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .3,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  geom_hline(
    yintercept = mp_wm_median,
    color = wm_color,
    linetype = "dashed",
    linewidth = hline_width
  ) +
  scale_color_manual(
    values = vg_clr_pal
  ) +
  scale_fill_manual(
    values = vg_clr_pal
  ) +
  scale_x_discrete(
    limits = rev(varied_graph_exp_structs)
  ) +
  scale_y_continuous(
    limits = c(0, 5.2),
    breaks = c(0, 1, 2, 3, 4, 5)
  ) +
  coord_flip() +
  theme(
    legend.position = "none"
  )
vg_multipath_fitness_plt
```

### Valley crossing - Valleys crossed

```{r}
vc_wm_median <- median(
  filter(
    max_org_data,
    (landscape == "Valley crossing") & (structure == "Well-mixed") & (experiment_set == "varied-graph")
  )$valleys_crossed
)

vg_valleys_plt <- max_org_data %>%
  filter(experiment_set == "varied-graph" & landscape == "Valley crossing") %>%
  mutate(
    structure = factor(
      structure,
      levels = varied_graph_exp_structs,
      labels = varied_graph_exp_structs
    )
  ) %>%
  ggplot(
    mapping = aes(
      x = structure,
      y = valleys_crossed,
      fill = structure
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping = aes(color = structure),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .3,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  geom_hline(
    yintercept = vc_wm_median,
    color = wm_color,
    linetype = "dashed",
    linewidth = hline_width
  ) +
  scale_color_manual(
    values = vg_clr_pal
  ) +
  scale_fill_manual(
    values = vg_clr_pal
  ) +
  scale_x_discrete(
    limits = rev(varied_graph_exp_structs)
  ) +
  coord_flip() +
  theme(
    legend.position = "none"
  )
vg_valleys_plt
```

## Squished toroid experiment set

### Smooth gradient - Time to maximum fitness

```{r}
st_clr_pal <- scale_color_bright()$palette(
  length(toroid_exp_structs)
)

# s <- color("bright")
# p <- plot_scheme(s(7), colours = TRUE)
# ggsave(
#   paste(plot_dir, "scheme.png", sep = "/"),
#   p
# )

st_smooth_grad_time_to_max_plt <- summary_gen_to_max %>%
  filter(experiment_set == "squished-toroid") %>%
  ggplot(
    mapping = aes(
      x = structure,
      y = proj_gen_max_mean,
      fill = structure,
      color = structure
    )
  ) +
  geom_col() +
  geom_linerange(
    aes(
      ymin = proj_gen_max_mean_ci_low,
      ymax = proj_gen_max_mean_ci_high
    ),
    color = "black",
    linewidth = 0.75,
    lineend = "round"
  ) +
  geom_hline(
    yintercept = max_generation,
    linetype = "solid",
    linewidth = 1,
    color = "darkgrey",
    alpha = hline_width
  ) +
  scale_x_discrete(
    breaks = toroid_exp_structs,
    limits = toroid_exp_structs
  ) +
  scale_color_manual(
    values = st_clr_pal
  ) +
  scale_fill_manual(
    values = st_clr_pal
  ) +
  coord_flip() +
  theme(
    legend.position = "none"
  )
st_smooth_grad_time_to_max_plt

# ggsave(
#   filename = paste(plot_dir, "st_smooth_grad_time_to_max.pdf", sep = "/"),
#   st_smooth_grad_time_to_max_plt
# )
```

### Multi-path exploration - Maximum fitness

```{r}
st_multipath_fitness_plt <- max_org_data %>%
  filter(experiment_set == "squished-toroid" & landscape == "Multipath") %>%
  mutate(
    structure = factor(
      structure,
      levels = toroid_exp_structs,
      labels = toroid_exp_structs
    )
  ) %>%
  ggplot(
    mapping = aes(
      x = structure,
      y = fitness,
      fill = structure
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping = aes(color = structure),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .3,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_color_manual(
    values = st_clr_pal
  ) +
  scale_fill_manual(
    values = st_clr_pal
  ) +
  scale_x_discrete(
    limits = toroid_exp_structs
  ) +
  scale_y_continuous(
    limits = c(0, 5.2),
    breaks = c(0, 1, 2, 3, 4, 5)
  ) +
  coord_flip() +
  theme(
    legend.position = "none"
  )
st_multipath_fitness_plt
```

### Valley crossing - Valleys crossed

```{r}
st_valleys_plt <- max_org_data %>%
  filter(experiment_set == "squished-toroid" & landscape == "Valley crossing") %>%
  mutate(
    structure = factor(
      structure,
      levels = toroid_exp_structs,
      labels = toroid_exp_structs
    )
  ) %>%
  ggplot(
    mapping = aes(
      x = structure,
      y = valleys_crossed,
      fill = structure
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping = aes(color = structure),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .3,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_color_manual(
    values = st_clr_pal
  ) +
  scale_fill_manual(
    values = st_clr_pal
  ) +
  scale_x_discrete(
    limits = toroid_exp_structs
  ) +
  scale_y_continuous(
    limits = c(0, 102),
    breaks = c(0, 25, 50, 75, 100)
  ) +
  coord_flip() +
  theme(
    legend.position = "none"
  )
st_valleys_plt
```

## Combined figures

### Plot grid for varied graph experiment set

Build plot grid for varied graph experiment set.

```{r}
vg_plt_grid <- plot_grid(
  vg_smooth_grad_time_to_max_plt +
    labs(
      x = "Graph type",
      y = "Gen. reached\nmax fitness"
    ) +
    theme(
      axis.text.y = element_text(angle = 30, hjust = 1),
      axis.text.x = element_text(angle = 30, hjust = 1),
      plot.margin = margin(1, 0, 0, 0, "cm")
    ),
  vg_multipath_fitness_plt +
    labs(
      y = "Final maximum\nfitness"
    ) +
    theme(
      axis.text.y = element_blank(),
      axis.title.y = element_blank(),
      axis.ticks.y = element_blank(),
      plot.margin = margin(1, 0, 0, 0, "cm")
    ),
  vg_valleys_plt +
    labs(
      y = "Valleys crossed"
    ) +
    theme(
      axis.text.y = element_blank(),
      axis.title.y = element_blank(),
      axis.ticks.y = element_blank(),
      plot.margin = margin(1, 0.2, 0, 0, "cm")
    ),
  nrow = 1,
  ncol = 3,
  align = "h",
  labels = c(
    "A) Gradient landscape",
    "B) Multi-path landscape",
    "C) Valley-crossing landscape"
  ),
  label_x = c(0.17, -0.4, -0.48),
  label_size = 11,
  rel_widths = c(1.5, 1, 1)
)
# vg_plt_grid

save_plot(
  filename = paste(plot_dir, "combined-vg-exp.pdf", sep = "/"),
  plot = vg_plt_grid,
  base_width = 8,
  base_height = 3.5,
  dpi = 600
)
```

### Plot grid for squished toroid experiment set

Build plot grid for squished toroid data

```{r}
st_plt_grid <- plot_grid(
  st_smooth_grad_time_to_max_plt +
    labs(
      x = "Lattice shape",
      y = "Gen. reached\nmax fitness"
    ) +
    theme(
      axis.text.y = element_text(angle = 30, hjust = 1),
      axis.text.x = element_text(angle = 30, hjust = 1),
      plot.margin = margin(1, 0, 0, 0, "cm")
    ),
  st_multipath_fitness_plt +
    labs(
      y = "Final maximum\nfitness"
    ) +
    theme(
      axis.text.y = element_blank(),
      axis.title.y = element_blank(),
      axis.ticks.y = element_blank(),
      # axis.text.x = element_text(angle = 0, hjust = 1),
      plot.margin = margin(1, 0, 0, 0, "cm")
    ),
  st_valleys_plt +
    labs(
      y = "Valleys crossed"
    ) +
    theme(
      axis.text.y = element_blank(),
      axis.title.y = element_blank(),
      axis.ticks.y = element_blank(),
      # axis.text.x = element_text(angle = 0, hjust = 1),
      plot.margin = margin(1, 0, 0, 0, "cm")
    ),
  nrow = 1,
  ncol = 3,
  align = "h",
  labels = c(
    "A) Gradient landscape",
    "B) Multi-path landscape",
    "C) Valley-crossing landscape"
  ),
  label_x = c(0.0, -0.36, -0.47),
  label_size = 11,
  rel_widths = c(1.5, 1, 1)
)

save_plot(
  filename = paste(plot_dir, "combined-st-exp.pdf", sep = "/"),
  plot = st_plt_grid,
  base_width = 8,
  base_height = 3.25,
  dpi = 600
)
```

### Full panel

```{r}

vg_plt_grid <- plot_grid(
  vg_smooth_grad_time_to_max_plt +
    labs(
      # title = "Smooth gradient",
      x = "Spatial Structure",
      y = "Gen. reached\nmax fitness"
    ) +
    theme(
      axis.text.y = element_text(angle = 30, hjust = 1),
      plot.margin = margin(1, 0, 0, 0, "cm")
    ),
  vg_multipath_fitness_plt +
    labs(
      # title = "Multipath",
      y = "Final max fitness"
    ) +
    theme(
      axis.text.y = element_blank(),
      axis.title.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.text.x = element_text(angle = 0, hjust = 1),
      plot.margin = margin(1, 0, 0, 0, "cm")
    ),
  vg_valleys_plt +
    labs(
      # title = "Valley crossing",
      y = "Valleys crossed"
    ) +
    theme(
      axis.text.y = element_blank(),
      axis.title.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.text.x = element_text(angle = 0, hjust = 1),
      plot.margin = margin(1, 0, 0, 0, "cm")
    ),
  nrow = 1,
  ncol = 3,
  align = "h",
  labels = c(
    "A) Gradient landscape",
    "B) Multipath landscape",
    "C) Valley crossing landscape"
  ),
  label_x = c(0.1, -0.4, -0.52),
  label_size = 12,
  rel_widths = c(1.5, 1, 1)
)

st_plt_grid <- plot_grid(
  st_smooth_grad_time_to_max_plt +
    labs(
      x = "Spatial Structure",
      y = "Gen. reached\nmax fitness"
    ) +
    theme(
      axis.text.y = element_text(angle = 30, hjust = 1),
      plot.margin = margin(1, 0, 0, 0, "cm")
    ),
  st_multipath_fitness_plt +
    labs(
      y = "Final max fitness"
    ) +
    theme(
      axis.text.y = element_blank(),
      axis.title.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.text.x = element_text(angle = 0, hjust = 1),
      plot.margin = margin(1, 0, 0, 0, "cm")
    ),
  st_valleys_plt +
    labs(
      y = "Valleys crossed"
    ) +
    theme(
      axis.text.y = element_blank(),
      axis.title.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.text.x = element_text(angle = 0, hjust = 1),
      plot.margin = margin(1, 0, 0, 0, "cm")
    ),
  nrow = 1,
  ncol = 3,
  align = "h",
  labels = c(
    "D) Gradient landscape",
    "E) Multipath landscape",
    "F) Valley crossing landscape"
  ),
  label_x = c(-0.05, -0.41, -0.52),
  label_size = 11,
  rel_widths = c(1.5, 1, 1)
)

all_exp_plt <- plot_grid(
  vg_plt_grid +
    theme(
      plot.margin = margin(1, 0, 0, 0, "cm")
    ),
  st_plt_grid +
    theme(
      plot.margin = margin(1, 0, 0, 1.7, "cm")
    ),
  labels = c(
    "Varied graph type experiments",
    "Varied lattice configuration experiments"
  ),
  label_size = 18,
  label_x = c(0.2, 0.1),
  nrow = 2,
  ncol = 1
)
all_exp_plt

save_plot(
  filename = paste(plot_dir, "simple-model-panel.pdf", sep = "/"),
  plot = all_exp_plt,
  base_width = 8,
  base_height = 8,
  dpi = 600
)
```