# Manuscript figures

## Analysis setup

```{r}
library(tidyverse)
library(cowplot)
library(RColorBrewer)
library(khroma)
library(rstatix)
library(knitr)
library(kableExtra)
library(boot)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

```{r}
varied_graph_experiment_slug <- "analysis_of_exploratory_data"
varied_graph_working_directory <- paste(
  "third-party",
  "evo_spatial_discoveries",
  "experiments",
  varied_graph_experiment_slug,
  sep = "/"
)

toroid_graph_experiment_slug <- "analysis_of_exploratory_toroid_data"
toroid_graph_working_directory <- paste(
  "third-party",
  "evo_spatial_discoveries",
  "experiments",
  toroid_graph_experiment_slug,
  sep = "/"
)

plt_dump_dir <- paste(
  "experiments",
  "mabe2-exps",
  "analyses",
  "plots",
  sep = "/"
)
```

```{r}
# Configure our default graphing theme
theme_set(theme_cowplot())
# Create a directory to store plots
plot_dir <- plt_dump_dir

dir.create(
  plot_dir,
  showWarnings = FALSE
)
```

## Load data

General parameters

```{r}
max_generation <- 100000
```

Load varied graph max organism data

```{r}
varied_graph_max_org_data_path <- paste(
  varied_graph_working_directory,
  "data",
  "combined_max_org_data.csv",
  sep = "/"
)
# Data file has time series
varied_graph_max_org_data_ts <- read_csv(varied_graph_max_org_data_path)
varied_graph_max_org_data_ts <- varied_graph_max_org_data_ts %>%
  mutate(
    landscape = as.factor(landscape),
    structure = as.factor(structure),
    experiment_set = "varied-graph"
  ) %>%
  mutate(
    valleys_crossed = case_when(
      landscape == "Valley crossing" ~ round(log(fitness, base = 1.5)),
      .default = 0
    )
  )
# Get tibble with just final generation
varied_graph_max_org_data <- varied_graph_max_org_data_ts %>%
  filter(generation == max_generation)
```

Load toroid max organism data

```{r}
toroid_graph_max_org_data_path <- paste(
  toroid_graph_working_directory,
  "data",
  "combined_max_org_data.csv",
  sep = "/"
)

# Data file has time series
toroid_graph_max_org_data_ts <- read_csv(toroid_graph_max_org_data_path)
toroid_graph_max_org_data_ts <- toroid_graph_max_org_data_ts %>%
  mutate(
    landscape = as.factor(landscape),
    structure = as.factor(structure),
    experiment_set = "squished-toroid"
  ) %>%
  mutate(
    valleys_crossed = case_when(
      landscape == "Valley crossing" ~ round(log(fitness, base = 1.5)),
      .default = 0
    )
  )
# Get tibble with just final generation
toroid_graph_max_org_data <- toroid_graph_max_org_data_ts %>%
  filter(generation == max_generation)
```

Combine toroid and varied graph data into single data frame.

```{r}
max_org_data_ts <- full_join(
  varied_graph_max_org_data_ts,
  toroid_graph_max_org_data_ts
)
max_org_data_ts <- max_org_data_ts %>%
  mutate(
    experiment_set = as.factor(experiment_set)
  )

structure_label_map <- list(
  "1_3600" = "1x3600",
  "15_240" = "15x240",
  "2_1800" = "2x1800",
  "3_1200" = "3x1200",
  "30_120" = "30x120",
  "4_900" = "4x900",
  "60_60" = "60x60",
  "well_mixed" = "Well-mixed",
  "comet_kite" = "Comet-kite",
  "windmill" = "Windmill",
  "clique_ring" = "Clique ring",
  "cycle" = "Cycle",
  "linear_chain" = "Linear chain",
  "lattice" = "Lattice",
  "random_waxman" = "Random Waxman",
  "star" = "Star",
  "wheel" = "Wheel"
)
toroid_exp_structs <- c(
  "1x3600",
  "15x240",
  "2x1800",
  "3x1200",
  "30x120",
  "4x900",
  "60x60"
)
varied_graph_exp_structs <- c(
  "Well-mixed",
  "Comet-kite",
  "Windmill",
  "Clique ring",
  "Cycle",
  "Linear chain",
  "Lattice",
  "Random Waxman",
  "Star",
  "Wheel"
)

structure_levels <- c(
  "1_3600",
  "2_1800",
  "3_1200",
  "4_900",
  "15_240",
  "30_120",
  "60_60",
  "well_mixed",
  "comet_kite",
  "windmill",
  "clique_ring",
  "cycle",
  "linear_chain",
  "lattice",
  "random_waxman",
  "star",
  "wheel"
)

structure_labels <- c()
for (struct_id in structure_levels) {
  structure_labels <- append(structure_labels, structure_label_map[[struct_id]])
}

max_org_data_ts <- max_org_data_ts %>%
  mutate(
    structure = factor(
      structure,
      levels = structure_levels,
      labels = structure_labels
    )
  )
# Get tibble with just final generation
max_org_data <- max_org_data_ts %>%
  filter(generation == max_generation)
```

Check that replicate count for each condition matches expectations.

```{r}
run_summary <- max_org_data %>%
  group_by(experiment_set, landscape, structure) %>%
  summarize(
    n = n()
  )
print(run_summary, n = 60)
```

Calculate time to max fitness for single-gradient landscape.

```{r}
# Find all rows with maximum fitness value, then get row with minimum generation,
#  then project out expected generation to max (for runs that didn't finish)
max_possible_fit = 50
time_to_max_single_gradient <- max_org_data_ts %>%
  filter(landscape == "Single gradient") %>%
  group_by(rep, structure, experiment_set) %>%
  slice_max(
    fitness,
    n = 1
  ) %>%
  slice_min(
    generation,
    n = 1
  ) %>%
  mutate(
    proj_gen_max = (max_possible_fit / fitness) * generation
  )
```

Calculate bootstrapped 99% confidence intervals for projected generation to maximum
fitness for single-gradient runs.

```{r}
# Define sample mean function
samplemean <- function(x, d) {
  return(mean(x[d]))
}

summary_gen_to_max <- tibble(
  experiment_set = character(),
  structure = character(),
  proj_gen_max_mean = double(),
  proj_gen_max_mean_ci_low = double(),
  proj_gen_max_mean_ci_high = double()
)
structures <- levels(time_to_max_single_gradient$structure)
experiment_sets <- levels(time_to_max_single_gradient$experiment_set)
for (exp_set in experiment_sets) {
  # cat("--------", exp_set, "--------")
  for (struct in structures) {
    # print(struct)
    s_data <- filter(
      time_to_max_single_gradient,
      (structure == struct) & (experiment_set == exp_set)
    )$proj_gen_max
    if (length(s_data) == 0) {
      next
    }
    boot_result <- boot(
      s_data,
      statistic = samplemean,
      R = 10000
    )
    result_ci <- boot.ci(boot_result, conf = 0.99, type = "perc")
    m <- result_ci$t0
    low <- result_ci$percent[4]
    high <- result_ci$percent[5]

    summary_gen_to_max <- summary_gen_to_max %>%
      add_row(
        experiment_set = exp_set,
        structure = struct,
        proj_gen_max_mean = m,
        proj_gen_max_mean_ci_low = low,
        proj_gen_max_mean_ci_high = high
      )
  }
}
summary_gen_to_max <- summary_gen_to_max %>%
  mutate(
    experiment_set = as.factor(experiment_set),
    structure = factor(
      structure,
      levels = structure_labels
    )
  )
```

## Varied graph type experiment set

### Smooth gradient - Time to maximum fitness

```{r}

# Identify median for well-mixed
wm_median <- median(
  filter(
    time_to_max_single_gradient,
    (structure == "Well-mixed") & (experiment_set == "varied-graph")
  )$proj_gen_max
)

vg_clr_pal <- scale_color_discreterainbow()$palette(
  length(varied_graph_exp_structs)
)

# color("discreterainbow")(length(varied_graph_exp_structs))
wm_color <- vg_clr_pal[1]

vg_smooth_grad_time_to_max_plt <- summary_gen_to_max %>%
  filter(experiment_set == "varied-graph") %>%
  ggplot(
    mapping = aes(
      x = structure,
      y = proj_gen_max_mean,
      fill = structure,
      color = structure
    )
  ) +
  # geom_point() +
  geom_col() +
  geom_linerange(
    aes(
      ymin = proj_gen_max_mean_ci_low,
      ymax = proj_gen_max_mean_ci_high
    ),
    color = "black",
    linewidth = 0.75,
    lineend = "round"
  ) +
  geom_hline(
    yintercept = max_generation,
    linetype = "dashed",
    linewidth = 1,
    color = "grey"
  ) +
  geom_hline(
    yintercept = wm_median,
    linetype = "dotted",
    color = wm_color,
    linewidth = 1
  ) +
  scale_x_discrete(
    breaks = rev(varied_graph_exp_structs),
    limits = rev(varied_graph_exp_structs)
  ) +
  scale_color_manual(
    values = vg_clr_pal
  ) +
  scale_fill_manual(
    values = vg_clr_pal
  ) +
  # scale_color_discreterainbow(
  #   breaks = varied_graph_exp_structs
  # ) +
  # scale_fill_discreterainbow(
  #   breaks = varied_graph_exp_structs
  # ) +
  coord_flip() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 30,
      hjust = 1
    )
  )
vg_smooth_grad_time_to_max_plt
```

### Multi-path exploration - Maximum fitness

```{r}
vg_multipath_fitness_plt <- max_org_data %>%
  filter(experiment_set == "varied-graph" & landscape == "Multipath") %>%
  mutate(
    structure = factor(
      structure,
      levels = varied_graph_exp_structs,
      labels = varied_graph_exp_structs
    )
  ) %>%
  ggplot(
    mapping = aes(
      x = structure,
      y = fitness,
      fill = structure
    )
  ) +
  geom_point(
    mapping = aes(color = structure),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .3,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_color_manual(
    values = vg_clr_pal
  ) +
  scale_fill_manual(
    values = vg_clr_pal
  ) +
  scale_x_discrete(
    limits = varied_graph_exp_structs
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 30,
      hjust = 1
    )
  )
vg_multipath_fitness_plt
```

### Valley crossing - Valleys crossed

```{r}

vg_valleys_plt <- max_org_data %>%
  filter(experiment_set == "varied-graph" & landscape == "Valley crossing") %>%
  mutate(
    structure = factor(
      structure,
      levels = varied_graph_exp_structs,
      labels = varied_graph_exp_structs
    )
  ) %>%
  ggplot(
    mapping = aes(
      x = structure,
      y = valleys_crossed,
      fill = structure
    )
  ) +
  geom_point(
    mapping = aes(color = structure),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .3,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_color_manual(
    values = vg_clr_pal
  ) +
  scale_fill_manual(
    values = vg_clr_pal
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 30,
      hjust = 1
    )
  )
vg_valleys_plt
```

## Squished toroid experiment set

## Combined figures