# 2024-11-14 Analysis Code


```{r}
library(tidyverse)
library(cowplot)
library(RColorBrewer)
library(khroma)
library(rstatix)
library(knitr)
library(kableExtra)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

```{r}
experiment_slug <- "2024-11-13-high-task-rewards"
working_directory <- paste(
  "experiments",
  experiment_slug,
  "analysis",
  sep = "/"
)
# Adjust working directory if being knitted for bookdown build.
# if (exists("bookdown")) {
#   if (bookdown) {
#     working_directory <- paste0(
#       bookdown_wd_prefix,
#       working_directory
#     )
#   }
# }
```

```{r}
# Configure our default graphing theme
theme_set(theme_cowplot())
# Create a directory to store plots
plot_dir <- paste(
  working_directory,
  "plots",
  sep = "/"
)

dir.create(
  plot_dir,
  showWarnings = FALSE
)
```

```{r}
# Load summary data from final update
data_path <- paste(
  working_directory,
  "data",
  "summary.csv",
  sep = "/"
)
data <- read_csv(data_path)

data <- data %>%
  mutate(
    graph_type = factor(
      graph_type,
      levels = c(
        "circular-ladder-graph",
        "connected-caveman",
        "cycle",
        "random-waxman",
        "well-mixed"
  
      )
    )
  )
```

```{r}
time_series_path <- paste(
  working_directory,
  "data",
  "time_series.csv",
  sep = "/"
)
time_series_data <- read_csv(time_series_path)

time_series_data <- time_series_data %>%
  mutate(
    graph_type = factor(
      graph_type,
      levels = c(
        "circular-ladder-graph",
        "connected-caveman",
        "cycle",
        "random-waxman",
        "well-mixed"
      )
    ),
    seed = as.factor(seed)
  )
```


```{r}

# Grab just completed rows
completed_runs_data <- data %>%
  filter(
    update == 200000
  )
```


## Number of tasks completed

```{r}
pop_tasks_total_plt <- ggplot(
  data = completed_runs_data,
  mapping = aes(
    x = graph_type,
    y = pop_task_total,
    fill = graph_type
  )
) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color = graph_type),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 30,
      hjust = 1
    )
  )

ggsave(
  filename = paste0(plot_dir, "/pop_tasks_total.pdf"),
  plot = pop_tasks_total_plt,
  width = 15,
  height = 10
)
```


```{r}
kruskal.test(
  formula = pop_task_total ~ graph_type,
  data = filter(completed_runs_data)
)

wc_results <- pairwise.wilcox.test(
  x = completed_runs_data$pop_task_total,
  g = completed_runs_data$graph_type,
  p.adjust.method	= "holm",
  exact = FALSE

)

pop_task_wc_table <- kbl(wc_results$p.value) %>% kable_styling()
save_kable(pop_task_wc_table, paste0(plot_dir, "/pop_task_wc_table.pdf"))
```

## Dominant tasks

```{r}
dom_tasks_total_plt <- ggplot(
  data = completed_runs_data,
  mapping = aes(
    x = graph_type,
    y = dom_task_total,
    fill = graph_type
  )
) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color = graph_type),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 30,
      hjust = 1
    )
  )

ggsave(
  filename = paste0(plot_dir, "/dom_tasks_total.pdf"),
  plot = dom_tasks_total_plt,
  width = 15,
  height = 10
)
```

```{r}
kruskal.test(
  formula = dom_task_total ~ graph_type,
  data = filter(completed_runs_data)
)
```

## MRCA Changes

```{r}
mrca_changes_plt <- ggplot(
  data = completed_runs_data,
  mapping = aes(
    x = graph_type,
    y = phylodiv_mrca_changes,
    fill = graph_type
  )
) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color = graph_type),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 30,
      hjust = 1
    )
  )

ggsave(
  filename = paste0(plot_dir, "/mrca_changes.pdf"),
  plot = mrca_changes_plt,
  width = 15,
  height = 10
)
```

## Average generation

```{r}
avg_generation_plt <- ggplot(
  data = completed_runs_data,
  mapping = aes(
    x = graph_type,
    y = time_average_generation,
    fill = graph_type
  )
) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color = graph_type),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 30,
      hjust = 1
    )
  )

ggsave(
  filename = paste0(plot_dir, "/avg_generation.pdf"),
  plot = avg_generation_plt,
  width = 15,
  height = 10
)
```


## Phylogenetic diversity
<!-- phylodiv_current_phylogenetic_diversity
phylodiv_diversity -->

```{r}
cur_phylo_div_plt <- ggplot(
  data = completed_runs_data,
  mapping = aes(
    x = graph_type,
    y = phylodiv_current_phylogenetic_diversity,
    fill = graph_type
  )
) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color = graph_type),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 30,
      hjust = 1
    )
  )

ggsave(
  filename = paste0(plot_dir, "/phylodiv_current_phylogenetic_diversity.pdf"),
  plot = cur_phylo_div_plt,
  width = 15,
  height = 10
)
```

```{r}
phylodiv_diversity_plt <- ggplot(
  data = completed_runs_data,
  mapping = aes(
    x = graph_type,
    y = phylodiv_diversity,
    fill = graph_type
  )
) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color = graph_type),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 30,
      hjust = 1
    )
  )

ggsave(
  filename = paste0(plot_dir, "/genotypic_diversity.pdf"),
  plot = phylodiv_diversity_plt,
  width = 15,
  height = 10
)
```

## Population task count over time

```{r}
pop_task_cnt_ts <- ggplot(
    data = time_series_data,
    mapping = aes(
      x = update,
      y = pop_task_total_tasks_done,
      color = graph_type,
      fill = graph_type
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  theme(legend.position = "bottom")

ggsave(
  plot = pop_task_cnt_ts,
  filename = paste0(
    working_directory,
    "/plots/pop_tasks_ts.pdf"
  ),
  width = 15,
  height = 10
)
```

```{r}
pop_task_cnt_ts <- ggplot(
    data = filter(time_series_data, update <= 10000),
    mapping = aes(
      x = update,
      y = pop_task_total_tasks_done,
      color = graph_type,
      fill = graph_type
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  theme(legend.position = "bottom")

ggsave(
  plot = pop_task_cnt_ts,
  filename = paste0(
    working_directory,
    "/plots/pop_tasks_ts_10k.pdf"
  ),
  width = 15,
  height = 10
)
```

## Diversity over time

```{r}
phylodiv_diversity_ts <- ggplot(
    data = time_series_data,
    mapping = aes(
      x = update,
      y = phylodiv_diversity,
      color = graph_type,
      fill = graph_type
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  theme(legend.position = "bottom")

ggsave(
  plot = phylodiv_diversity_ts,
  filename = paste0(
    working_directory,
    "/plots/phylodiv_diversity_ts.pdf"
  ),
  width = 15,
  height = 10
)
```

```{r}
phylodiv_current_phylogenetic_diversity_ts <- ggplot(
    data = time_series_data,
    mapping = aes(
      x = update,
      y = phylodiv_current_phylogenetic_diversity,
      color = graph_type,
      fill = graph_type
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  theme(legend.position = "bottom")

ggsave(
  plot = phylodiv_current_phylogenetic_diversity_ts,
  filename = paste0(
    working_directory,
    "/plots/phylodiv_current_phylogenetic_diversity_ts.pdf"
  ),
  width = 15,
  height = 10
)
```

## Average generation over time

```{r}
time_average_generation_ts <- ggplot(
    data = time_series_data,
    mapping = aes(
      x = update,
      y = time_average_generation,
      color = graph_type,
      fill = graph_type
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  theme(legend.position = "bottom")

ggsave(
  plot = time_average_generation_ts,
  filename = paste0(
    working_directory,
    "/plots/time_average_generation_ts.pdf"
  ),
  width = 15,
  height = 10
)
```

## MRCA changes over time

```{r}
mrca_changes_ts <- ggplot(
    data = time_series_data,
    mapping = aes(
      x = update,
      y = phylodiv_mrca_changes,
      color = graph_type,
      fill = graph_type
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  theme(legend.position = "bottom")

ggsave(
  plot = mrca_changes_ts,
  filename = paste0(
    working_directory,
    "/plots/mrca_changes_ts.pdf"
  ),
  width = 15,
  height = 10
)
```

```{r}
mrca_changes_ts <- ggplot(
    data = filter(
      time_series_data,
      update >= 150000 & update <= 200000
    ),
    mapping = aes(
      x = update,
      y = phylodiv_mrca_changes,
      color = graph_type,
      fill = graph_type
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  theme(legend.position = "bottom")

ggsave(
  plot = mrca_changes_ts,
  filename = paste0(
    working_directory,
    "/plots/mrca_changes_ts_mod.pdf"
  ),
  width = 15,
  height = 10
)
```

## Average phylo depth over time

```{r}
phylo_ave_depth_ts <- ggplot(
    data = time_series_data,
    mapping = aes(
      x = update,
      y = phylodiv_ave_depth,
      color = graph_type,
      fill = graph_type
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  theme(legend.position = "bottom")

ggsave(
  plot = phylo_ave_depth_ts,
  filename = paste0(
    working_directory,
    "/plots/phylo_ave_depth_ts.pdf"
  ),
  width = 15,
  height = 10
)
```
## Graph location info

Analyze graph_birth_info_annotated.csv

```{r}
# Load summary data from final update
graph_loc_data_path <- paste(
  working_directory,
  "data",
  "graph_birth_info_annotated.csv",
  sep = "/"
)
graph_loc_data <- read_csv(graph_loc_data_path)

graph_loc_data <- graph_loc_data %>%
  mutate(
    graph_type = factor(
      graph_type,
      levels = c(
        "circular-ladder-graph",
        "connected-caveman",
        "cycle",
        "random-waxman",
        "well-mixed"
      )
    ),
    seed = as.factor(seed)
  )
```

Summarize by seed
```{r}
graph_loc_data_summary <- graph_loc_data %>%
  group_by(seed, graph_type) %>%
  summarize(
    births_var = var(births),
    births_total = sum(births),
    task_apps_total = sum(task_appearances),
    task_apps_var = var(task_appearances)
  ) %>%
  ungroup()
```

### Total birth Counts

```{r}
birth_counts_total_plt <- ggplot(
  data = graph_loc_data_summary,
  mapping = aes(
    x = graph_type,
    y = births_total,
    fill = graph_type
  )
) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color = graph_type),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 30,
      hjust = 1
    )
  )

ggsave(
  filename = paste0(plot_dir, "/birth_counts_total.pdf"),
  plot = birth_counts_total_plt,
  width = 15,
  height = 10
)
```

### Variance birth Counts

```{r}
birth_counts_var_plt <- ggplot(
  data = graph_loc_data_summary,
  mapping = aes(
    x = graph_type,
    y = births_var,
    fill = graph_type
  )
) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color = graph_type),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 30,
      hjust = 1
    )
  )

ggsave(
  filename = paste0(plot_dir, "/birth_counts_var.pdf"),
  plot = birth_counts_var_plt,
  width = 15,
  height = 10
)
```

### Task appearances total

```{r}
task_apps_total_plt <- ggplot(
  data = graph_loc_data_summary,
  mapping = aes(
    x = graph_type,
    y = task_apps_total,
    fill = graph_type
  )
) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color = graph_type),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 30,
      hjust = 1
    )
  )

ggsave(
  filename = paste0(plot_dir, "/task_apps_total.pdf"),
  plot = task_apps_total_plt,
  width = 15,
  height = 10
)
```

```{r}
kruskal.test(
  formula = task_apps_total ~ graph_type,
  data = graph_loc_data_summary
)

wc_results <- pairwise.wilcox.test(
  x = graph_loc_data_summary$task_apps_total,
  g = graph_loc_data_summary$graph_type,
  p.adjust.method	= "holm",
  exact = FALSE

)
wc_results
```


## Moran's I results

```{r}
# Load summary data from final update
morans_i_data_path <- paste(
  working_directory,
  "data",
  "morans_i.csv",
  sep = "/"
)
morans_i_data <- read_csv(morans_i_data_path)

morans_i_data <- morans_i_data %>%
  mutate(
    graph_type = str_split_i(
      graph_file,
      pattern = ".mat",
      1
    )
  ) %>%
  mutate(
    graph_type = factor(
      graph_type,
      levels = c(
        "circular-ladder-graph",
        "connected-caveman",
        "cycle",
        "random-waxman",
        "well-mixed"
      )
    ),
    seed = as.factor(seed)
  )
```

### Clustered task appearances

Summarize statistically significant runs where I > 0.

```{r}
# Identify number of runs where distribution of task appearances is more
# clustered than we would expect by chance.
clustered_counts <- morans_i_data %>%
  filter(
    (task_morans_i > 0) & (task_p_val <= 0.05)
  ) %>%
  group_by(graph_type) %>%
  summarize(
    n = n()
  )
```

```{r}
tasks_clustered_plt <- clustered_counts %>%
  ggplot(
    aes(
      x = graph_type,
      y = n,
      color = graph_type,
      fill = graph_type
    )
  ) +
  geom_col() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 30,
      hjust = 1
    )
  )

ggsave(
  filename = paste0(plot_dir, "/tasks_clustered_plt.pdf"),
  plot = tasks_clustered_plt,
  width = 15,
  height = 10
)
```

```{r}
tasks_clustered_i_vals_plt <- morans_i_data %>%
  filter((task_morans_i > 0) & (task_p_val <= 0.05)) %>%
  ggplot(
    mapping = aes(
      x = graph_type,
      y = task_morans_i,
      fill = graph_type
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color = graph_type),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 30,
      hjust = 1
    )
  )

ggsave(
  filename = paste0(plot_dir, "/tasks_clustered_i_vals_plt.pdf"),
  plot = tasks_clustered_i_vals_plt,
  width = 15,
  height = 10
)
```

### Clustered birth counts

```{r}
births_clustered_plot <- morans_i_data %>%
  filter((birth_morans_i > 0) & (birth_p_val <= 0.05)) %>%
  ggplot(
    aes(
      x = graph_type,
      color = graph_type,
      fill = graph_type
    )
  ) +
  geom_bar() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 30,
      hjust = 1
    )
  )

ggsave(
  filename = paste0(plot_dir, "/births_clustered_plot.pdf"),
  plot = births_clustered_plot,
  width = 15,
  height = 10
)
```

```{r}
birth_clustered_i_vals_plt <- morans_i_data %>%
  filter((birth_morans_i > 0) & (birth_p_val <= 0.05)) %>%
  ggplot(
    mapping = aes(
      x = graph_type,
      y = birth_morans_i,
      fill = graph_type
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color = graph_type),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 30,
      hjust = 1
    )
  )

ggsave(
  filename = paste0(plot_dir, "/birth_clustered_i_vals_plt.pdf"),
  plot = birth_clustered_i_vals_plt,
  width = 15,
  height = 10
)
```

## End